Community Data Science meets Digital Heritage: Controversies about World Heritage Sites in the Online Community of Wikipedia

### 2019-05-13 code/paper.Rmd
time series and inscription years

Load the data

@@

## Some sort of clustering

@@

@@

Here's a map of countries showing which has the most WP pages for WH sites

@@

Map of the ratio of WH sites with WP pages, per country

@@

@@

Scatterplot showing the WP sites that get the most attention

@@

Edits per 1k words, as a measure of edit density

@@

Number of unique editors per page

@@

Number of unique editors per 1k words per page

@@

Size of edits, absolute and raw

@@

How many edits made by bots? Proportion of edits per page made by bots

@@

Which WH pages are most edited by bots?

@@

which bots are most active on WH pages vs random pages?
Cydebot: category edits
Smackbot: dates in templates
ClueBot NG: vandalism

@@

So about vandalism? do WH sites get more than random ones?

@@

Which WH pages have the most vandalism?

@@

Reverts

@@

Which articles have the most reverts?

@@

Is there a correlation between reverts and vandalism?

@@

What's the typical distribution of talk page length?

@@

which WH pages have the longest talk pages?

@@

Talk page topic modelling

@@

The probability that each document is generated from each topic.

@@

we may want to know which topics are associated with each talk page?

@@

Topic correlations

@@

IP addresses

@@

How many IP edits on each page?

@@

Overall location of IP edits for all random pages

@@

What's the geographic diversity of edits per random page?

@@

@@

time between edits

@@

Page ages: are WH pages younger than the typical WP page? What's their relationship to their inscription year?

@@

) %>% 
  ggplot() +
geom_density(aes(as.numeric(edit_span), 
                 fill = source),
                 alpha = 0.3) +
  # scale_x_log10() +
  scale_fill_viridis_d() +  
  theme_minimal(base_size = 12) + 
  xlab("Age of Wikipedia page (days)")

ggsave("figures/wh_wikpedia_pages_ages.png")

Time between inscription year and first edit

@@

time series of edits: breakouts relating to events? monthly seems like a good level of aggregation. What can we do for spike detection?

```{r}
randoms_revision_history_page_details_ip_location %>% 
  group_by(page) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(page, the_m) %>% 
  tally() %>% 
  filter(n() > 130) %>%   # explore this value
  ggplot() +
  geom_line(aes(the_m,
             n)) +
  facet_wrap( ~ page, scales = "free")

Spike detection

@@

```{r}
# devtools::install_github("hrbrmstr/AnomalyDetection")
library(AnomalyDetection)
randoms_revision_history_page_details_ip_location_monthly <- 
randoms_revision_history_page_details_ip_location %>% 
  group_by(page) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(page, the_m) %>% 
  tally() %>% 
  filter(n() > 150) # explore this value

>>>>

## Load the data

@@

## Overall clusters

@@

## More maps

@@

## Here's a map of countries showing which has the most WP pages for WH sites

@@

## Map of the ratio of WH sites with WP pages, per country

@@

# basic qualities scatterplot

@@

## Scatterplot showing the WP sites that get the most attention

@@

## Edits per 1k words, as a measure of edit density

@@

## Number of unique editors per page

@@

## Number of unique editors per 1k words per page

@@

## Size of edits, absolute and raw

@@

## How many edits made by bots? Proportion of edits per page made by bots

@@

## Which WH pages are most edited by bots?

@@

## which bots are most active on WH pages vs random pages?
Cydebot: category edits
Smackbot: dates in templates
ClueBot NG: vandalism

@@

## So about vandalism? do WH sites get more than random ones?

@@

## Which WH pages have the most vandalism?

@@

## Reverts

@@

## Which articles have the most reverts?

@@

## Is there a correlation between reverts and vandalism?

@@

## What's the typical distribution of talk page length?

@@

## which WH pages have the longest talk pages?

@@

## Talk page topic modelling

@@

## The probability that each document is generated from each topic.

@@

## we may want to know which topics are associated with each talk page?

@@

## Topic correlations

@@

## IP addresses

@@

## How many IP edits on each page?

@@

## Overall location of IP edits for all random pages

@@

## What's the geographic diversity of edits per random page?

@@

## Circular chord plot

@@

## time between edits

@@

## Page ages: are WH pages younger than the typical WP page? What's their relationship to their inscription year?

@@

) %>% 
  ggplot() +
geom_density(aes(as.numeric(edit_span / 365), 
                 fill = source),
                 alpha = 0.3) +
  geom_vline(xintercept = year(Sys.Date()) - 2001, 
             colour = "red") +
  annotate("text", x = 15, y = 0.2, label = "start of Wikipedia → ") +
  scale_fill_viridis_d() +  
  theme_minimal(base_size = 12) + 
  xlab("Age of Wikipedia page (years)")

ggsave("figures/wh_wikpedia_pages_ages.png")
```

## Time between inscription year and first edit

@@

`all_the_data_to_load_quickly.RData` is up to here

```{r}
load("data/all_the_data_to_load_quickly.RData")

# remove some big items
rm(list = ls(pattern = "random"))
library(tidyverse)
library(lubridate)
```

## time series of edits: breakouts relating to events? monthly seems like a good level of aggregation. What can we do for spike detection?

```{r}
wh_pages_edit_time_series <- 
wh_pages_talk_page_length %>% 
  unnest(revision_history_page_details) %>% 
  filter(Year_num >= 2001) %>% # only sites inscribed after WP started
  group_by(Site) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(Site, the_m) %>% 
  tally() %>% 
  filter(n() > 100) %>% # explore this value
  left_join(wh_pages_talk_page_length) %>% 
  mutate(vline_year = ifelse(Year_num >= 2001, Year_num, NA)) %>% 
  mutate(vline_year = ymd_hms(ISOdatetime(vline_year, 1, 1, 1, 0, 0)))

ggplot(wh_pages_edit_time_series) +
  geom_vline(aes(xintercept = vline_year), colour = "red") +
  geom_line(aes(the_m,
             n)) +
  facet_wrap( ~ str_wrap(Site, 30), scales = "free") +
  theme_minimal(base_size = 6)

ggsave("figures/wh_wikpedia_pages_edit_time_series.png", h = 9, w = 12)
```

## Spike detection

@@

```{r}
# devtools::install_github("hrbrmstr/AnomalyDetection")
library(AnomalyDetection)
randoms_revision_history_page_details_ip_location_monthly <- 
randoms_revision_history_page_details_ip_location %>% 
  group_by(page) %>% 
  mutate(the_d = floor_date(rh_date, "day")) %>% 
    mutate(the_m = floor_date(rh_date, "month")) %>% 
  group_by(page, the_m) %>% 
  tally() %>% 
  filter(n() > 200) # explore this value
